Summary stats for paper

### 2021-05-27 analysis/match_data_sources.Rmd
updated figs and stats with newly run farms

all_farms <- list.files("data/temp_data/temp_marine_final", pattern = "rds$", full.names = TRUE) %>% 
  map_df(., readRDS) %>% 
  rename(iso3c = iso3) %>% 
  left_join(all_farms_to_place) %>%
  mutate(source = "modeled") %>%
  mutate(estimated = ifelse(iso3c == "IND" & type == "bivalve", 0, estimated)) %>% # fix india mistake
  rbind(data %>% mutate(source = "real", estimated = 0)) %>%
  mutate(data_type = case_when(
    source == "real" ~ "known location", 
    source == "modeled" & estimated == 0 ~ "modeled location, known number of farms",
    source == "modeled" & estimated == 1 ~ "modeled location, estimated number of farms"
  ))
summary(all_farms)
```

@@

@@

summary(all_farms_sources_final)
unique(all_farms_sources_final$data_year)

>>>>

all_farms <- readRDS("data/all_modeled_marine_farms.rds") %>%
  left_join(all_farms_to_place) %>%
    mutate(data_type = case_when(
    source == "real" ~ "known location", 
    source == "modeled" & estimated == 0 ~ "modeled location, known number of farms",
    source == "modeled" & estimated == 1 ~ "modeled location, estimated number of farms"
  ))

summary(all_farms)

test <- all_farms %>%
  st_drop_geometry() %>%
  group_by(data_type) %>% 
  summarise(n()) ## perfect!!

@@

# check <- USA_salmon_fix %>%
#   st_drop_geometry() %>%
#   group_by(data_source) %>%
#   summarise(n()) ## numbers are correct

@@

summary(all_farms_sources_final)
unique(all_farms_sources_final$data_year)
unique(sort(all_farms_sources$country))
unique(sort(all_farms_sources$iso3c)) ## 73 countries - perfect!

### 2021-05-27 analysis/paper_figs.Rmd
did some checking for paper
updated figs and stats with newly run farms

test <- all_farms_sf %>%
  filter(country == "United States of America")
mapview(test) ## some being placed around great lakes... need to fix that

@@

@@

>>>>

test <- all_farms_sf %>%
  filter(country == "United States of America")
mapview(test)

test <- all_farms_sf %>%
  filter(iso3c == "HRV")
mapview(test) ## one weird MFG from ASC

test <- all_farms_sf %>%
  filter(iso3c == "TUR")
mapview(test) ## one weird MFG from ASC.. i think these are fine. Justifiable since it is from the ASC...

@@

## Radial bar chart of estimation type per country

```{r}
library(tidyverse)
library(viridis)
library(hrbrthemes)
library(RColorBrewer)
library(colorspace)

rank_rgn <- all_farms %>%
  mutate(country = ifelse(country == "United Kingdom (Scotland) ", "United Kingdom", country)) %>%
  group_by(country) %>%
  summarise(num_farms_total = n()) %>%
  ungroup() %>%
  arrange(num_farms_total)

## Stuff for both plots
# i think that this would look better with each stressor having its own color that is paired with ocean/land
# 
# pick 4 distinct colors and then create a pair with one being more brown or green and the other being more blue

source_name <- data.frame(source = c("known location", "modeled location, known number of farms", "modeled location, estimated number of farms"))

type_1 <- "forestgreen"
type_1 <- lighten(type_1, amount = -0.05)
type_2 <- "orange1"
type_2 <- lighten(type_2, amount = -0.05)
type_3 <- "red3"
type_3 <- lighten(type_3, amount = -0.05)

gc_palette <- c(type_1, type_2, type_3)
scales::show_col(gc_palette, n = 3)

pal_df <- as_tibble(gc_palette) %>% 
  rename(fill_value = value) %>% 
  mutate(source = case_when(fill_value == type_1 ~ "known location",
                            fill_value == type_2 ~ "modeled location, known number of farms",
                            fill_value == type_3 ~ "modeled location, estimated number of farms"))

## circle plot

cutoff_value <-  1000

by_country <- all_farms %>%
  mutate(country = ifelse(country == "United Kingdom (Scotland) ", "United Kingdom", country)) %>%
  group_by(country, data_type) %>%
  summarise(num_farms = n()) %>%
  ungroup() %>%
  group_by(country) %>%
  mutate(num_farms_total = sum(num_farms)) %>%
  ungroup() %>%
  arrange(num_farms_total) %>%
  filter(num_farms_total < cutoff_value)

##grab the highest country to highlight red and id where the circle ends and then side bar chart begins
highest_country <- by_country[dim(by_country)[1], "country"]
farms_highest_country <- filter(by_country, country %in% highest_country$country) #used below

# adjust the rank_rgn-adjust dataframe to match above
rank_rgn_adj <- rank_rgn %>%
  filter(country %in% by_country$country)

# add empty space to create space in the circleplot for axes
 empty_bar <- 5
 to_add <- data.frame(matrix(NA, empty_bar, ncol(rank_rgn_adj)) )
 colnames(to_add) = colnames(rank_rgn_adj)
 to_add$country <- as.character(1:empty_bar)
 rank_rgn_adj  <- rbind(to_add, rank_rgn_adj)

 # some code to orient the country labels
sequence_length = length(unique(rank_rgn_adj$country))
first_sequence = c(1:(sequence_length%/%2)) 
second_sequence = c((sequence_length%/%2+1):sequence_length) 
first_angles = c(90 - 180/length(first_sequence) * first_sequence)
second_angles = c(-90 - 180/length(second_sequence) * second_sequence)
rank_rgn_adj$angle <- c(first_angles, second_angles)
rank_rgn_adj$hjust <- c(rep(0, length(first_sequence)), 
                    rep(1, length(second_sequence)))

# color for region labels
rank_rgn_adj <- rank_rgn_adj %>%
  mutate(country = factor(country, unique(country))) %>%
  mutate(color = "black") %>%
  mutate(color = ifelse(is.na(num_farms_total), "white", color)) %>%
  mutate(color = ifelse(country %in% highest_country$country, "red", color))

## make the plot

circle_theme <- theme(axis.line=element_blank(),
                      axis.text.y=element_blank(),
                      axis.ticks=element_blank(),
                      axis.title.x=element_blank(),
                      axis.title.y=element_blank(),
                      legend.position="none",
                      panel.background=element_blank(),
                      panel.border=element_blank(),
                      panel.grid.major=element_blank(),
                      panel.grid.minor=element_blank(),
                      plot.background=element_blank(),
                      axis.text.x = element_blank())

color_order <- c(type_1, type_2, type_3)

pressures_adj_circle_plot <- by_country %>% 
  mutate(prop_of_global_stacked = abs(num_farms)) %>% 
  left_join(pal_df, by = c("data_type" = "source")) %>% 
  mutate(fill_value = factor(fill_value, levels = color_order))

 ggplot(data=pressures_adj_circle_plot, aes(x=country, y=prop_of_global_stacked, fill = fill_value)) +

  geom_bar(stat="identity") +

  # ## this adjust the inner circle size (ymin) and outer circle size (ymax), which can be expanded to include names
   geom_errorbar(aes(
       x     = 1,
        ymin = -0.01,   # ymin  = -0.03, this is for when ocean bars stack inwards -0.01 for when not
       ymax  = 0.035),
       alpha = 0) +
      # Country name text placement
     geom_text(
       data = rank_rgn_adj,
       aes(x     = country,
           y     = max(by_country$num_farms_total, na.rm=TRUE),
           label = country,
           angle = angle,
           hjust = hjust),
       inherit.aes = FALSE,
       fontface = "bold",
       alpha = 0.9,
       size = 4,
       color = rank_rgn_adj$color) +
    # 
   # georegion names
    # geom_text(
    #   data=rgn_shift,
    #   aes(y = 0.02, x = name_x, label = georegion),
    #   # aes(x=name_x,
    #   #     y=name_y,
    #   #     label=georegion),
    #   inherit.aes=FALSE, size=7) +
  #  
  # # # Scale bar circles
  # ## This adds the 0 line
      geom_segment(
        x = 0, 
         y = 0, 
         xend = dim(rank_rgn_adj)[1]+1, 
         yend = 0, 
         fill = NULL,
        alpha = 1, color = "white", size = 0.1) + #color = "black for internal
  #    
  #  # Scale bar annotation, this adds the y-axis numbers
      annotate("text",
        x     = c(3, 3, 3),
        y     = c(0, 0.0175, 0.035),
        label = c(0, 0.0175, 0.035),
        color = "darkgray", angle = -4, size = 4) +
  #     
  scale_color_identity() +
   scale_fill_identity() +
   # scale_fill_manual(values= rev(myPalette), c("Land: Disturbance", "Land: GHG", "Land: Water", "Land: Nutrient",
   #                                    "Ocean: Disturbance", "Ocean: GHG", "Ocean: Water", "Ocean: Nutrient")) +
  coord_polar() +
 circle_theme

@@

test_data <- all_farms %>%
  filter(data_type != "known location")

## create first plot
total_map <- ggplot(data = world) +
  geom_sf(fill = "white", lwd = 0.1) + 
  geom_point(data = test_data, aes(x = X, y = Y, col = data_type), size = 0.1, alpha = 0.7) +
#  coord_sf(xlim = c(-180, 180), ylim = c(-58, 73), expand = FALSE, label_axes = list(top = "E", left = "N")) +
  coord_sf(xlim = c(-180, 180), ylim = c(-58, 73), expand = FALSE) +## or could make the latitude distances even i.e. (-72, 72)
  scale_x_continuous(breaks = c(-180, -135, -90, -45, 0, 45, 90, 135, 180), position = "top") +
  scale_y_continuous(breaks = c(-90, -45, 0, 45, 90)) + 
  theme_bw() + 
  theme(axis.title.x = element_blank(),
          axis.title.y = element_blank(),
        legend.position = "bottom", 
        legend.title=element_blank()) +
  guides(color = guide_legend(override.aes = list(size = 3, alpha = 0.9), nrow=2 )) +
  scale_color_manual(values = c("forestgreen", "orange1", "red3"), 
                    breaks = c("known location", "modeled location, known number of farms", "modeled location, estimated number of farms"), labels = c("known location \n n = 9,750", "modeled location, known number of farms \n n = 10,519", "modeled locations, estimated number of farms \n n = 75,073"))
    #  geom_vline(xintercept = 135) + # check to make sure it lines up
  #geom_hline(yintercept = 0) # check to make sure it lines up

ggsave("analysis/figs/map_fig_test.jpg")

### 2021-05-27 analysis/paper_stats.Rmd
updated figs and stats with newly run farms

library(here)
library(dplyr)
library(sf)
library(raster)
library(mapview)
library(tidyverse)
library(RCurl)
library(mosaic)
library(googlesheets4)
library(janitor)

@@

# Global regions

@@

# this file includes ocean regions:
gadm <- raster("/home/shares/food-systems/Food_footprint/_raw_data/Country_regions/gadm36mol_pyfill_10km.tif")

gadm_latlon <- raster::projectRaster(gadm, food_raster, method="ngb", over=TRUE)
plot(gadm_latlon)

gadm_codes <- read.csv("/home/shares/food-systems/Food_footprint/_raw_data/Country_regions/gadm_reg_codes.csv") %>% 
  mutate(reg_id = as.numeric(reg_id))
```

# Read in farms data and convert to sf

```{r}
all_farms_sf <- read_csv("all_marine_aquaculture_farms_sources.csv") %>%
    st_as_sf(., 
           coords = c("X", "Y"),
           crs = crs(food_raster), 
           agr = "constant")

all_farms <- all_farms_sf %>%
  st_drop_geometry()

@@

# Chile, Norway, Peru, the United States, Canada and Ireland provided a substantial number of known locations (>1000).
top_known <- all_farms %>%
  filter(data_type == "known location") %>%
  group_by(iso3c) %>%
    summarise(n_obs = n())

@@

by_species_data_type <- all_farms %>%
  group_by(species, data_type) %>%
  summarise(n_obs = n()) %>%
  mutate(total_obs = sum(n_obs)) %>% 
  ungroup() %>%
  mutate(pct = n_obs/total_obs*100)
```

Check some of the data
```{r}
by_type <- all_farms %>%
  group_by(species) %>%
  summarise(n_obs = n())

>>>>

library(here)
library(dplyr)
library(sf)
library(raster)
library(mapview)
library(tidyverse)
library(RCurl)
library(mosaic)
library(googlesheets4)
library(janitor)
library(countrycode)

@@

# Read in farms data

@@

all_farms <- read_csv("all_marine_aquaculture_farms_sources.csv")

all_farms_sf <- all_farms %>%
  st_as_sf(coords =c("X", "Y"))

# test <- all_farms_sf %>%
#   filter(iso3c == "USA")
# mapview(test)
```

@@

# Chile, Norway, and Ireland provided a substantial number of known locations (>1000).
top_known <- all_farms %>%
  filter(data_type == "known location") %>%
  group_by(iso3c) %>%
    summarise(n_obs = n())

@@

by_species_data_type <- all_farms %>%
  group_by(species, data_type) %>%
  summarise(n_obs = n()) %>%
  mutate(total_obs = sum(n_obs)) %>% 
  ungroup() %>%
  mutate(pct = n_obs/total_obs*100)

## How much mariculture production was included in our assessment?

fao <- read_csv("data/fao_aquaculture_clean.csv")
mar_sp_list <- read_csv("https://raw.githubusercontent.com/OHI-Science/ohiprep_v2020/gh-pages/globalprep/mar/v2020/raw/species_list.csv")

fw <- fao %>%
  dplyr::filter(environment == "Freshwater") %>%
  group_by(species) %>%
  summarize(total = sum(value)) %>%
  ungroup() %>%
  arrange(total)

data.frame(fw)

test <- fao %>%
  filter(year == 2017) %>%
  left_join(mar_sp_list, by = c("species" = "FAO_name")) %>%
  mutate(environment2 = ifelse(Taxon_code == "AL", "seaweed", environment)) %>%
  mutate(environment = ifelse(is.na(environment2), environment, environment2)) %>%
  group_by(environment, year) %>%
  summarise(prod = sum(value, na.rm = TRUE)) %>%
  ungroup() %>%
  mutate(total = sum(prod)) %>%
  mutate(pct = prod/total)

## total marine animal production is: 8090542 + 22920871 = 31011413

## we include: 29623515
prod_total <- read_csv("/home/shares/food-systems/Food_footprint/all_food_systems/dataprep/aquaculture/marine/production_extracted_final.csv")

sum(prod_total$total_tonnes) # 29623515

29623515/31011413 # 0.9552456
```

Countries with comprehensive farm locations 
```{r}

all_farms <- read_csv("all_marine_aquaculture_farms_sources.csv")

unique(sort(all_farms$country))

country_spp_combos <- all_farms %>%
  distinct(country, species)

## read in each of the comprehensive locations

bivalve_comp <- read_csv("marine/bivalve/bivalve_farms/output/bivalve_comp_tonnes_per_farm.csv")

shrimp_comp <- read_csv("marine/shrimp/shrimp_farms/output/shrimp_comp_tonnes_per_farm.csv")

salmon_comp <- read_csv("marine/salmon/salmon_farms/output/salmon_comp_tonnes_per_farm.csv")

tuna_comp <- read_csv("marine/tuna/tuna_farms/output/tuna_comp_tonnes_per_farm.csv")

finfish_comp <- read_csv("marine/marine_fish_general/finfish_farms/output/finfish_comp_tonnes_per_farm.csv")

all_comp_farms <- rbind(finfish_comp, bivalve_comp, salmon_comp, tuna_comp, shrimp_comp) %>%
    mutate(Species = case_when(
      type == "salmon" ~ "Salmonids", 
      type == "bivalve" ~ "Unfed or algae fed shellfish",
      type == "marine_fish_general" ~ "General Marine Fish", 
      type == "shrimp" ~ "Shrimp", 
      type == "tuna" ~ "Tuna"
    )) %>%
  mutate(tonnes_per_farm = round(tonnes_per_farm, 2)) %>%
  dplyr::select("Country" = "country", Species, "Number of farms" = "n_farms", "Tonnes of production" = "fao_tonnes_production", "Tonnes per farm" = "tonnes_per_farm") %>%
  arrange(Country)

unique(all_comp_farms$Country)

write.csv(all_comp_farms, "analysis/output/all_comp_farms.csv", row.names = FALSE)

comp_across_all <- all_comp_farms %>%
  full_join(country_spp_combos, by = c("Country" = "country", "Species" = "species")) %>%
  mutate(Country = case_when(
    Country == "Iran" ~ "Iran (Islamic Rep. of)", 
    Country == "Hong Kong " ~ "China, Hong Kong SAR",
    TRUE ~ Country
  )) %>%
  filter(Country %in% all_comp_farms$Country) %>%
  group_by(Country) %>%
  summarise(n_obs = sum(`Number of farms`), 
            n_species = n(),
            tonnes = sum(`Tonnes of production`)) %>%
  filter(!is.na(n_obs))

sum(comp_across_all$tonnes) # account for 4328484 tonnes of production

4328484/29623515 # these account for ~15% of production tonnage

## these are the countries that are comprehensive across all species groups they report
cat(comp_across_all$Country, sep = ", ")

# tuna_spp <- read_csv("marine/tuna/tuna_farms/data/fao_mariculture_tuna_taxa_production.csv")
```

Score the different taxa maps for the global food paper:

Score | Spatial Extent | Spatial Resolution | Temporal specificity | System specificity | System comprehensiveness
-- | -- | -- | -- | -- | --
5 | Represents entire spatial extent of the food system. | <= 5 arc minutes | Data is specific to 2017 | The data is reported to the level of our specific food type (e.g. cow mixed milk) | Data represents all foods within a system
4 | Need to gapfill regions with relatively little production (e.g. Aland islands) | sub-national or raster with cells larger than 5-arc minutes | Data range includes 2017 and is likely to represent 2017 | Generalized to animal type, but not both rearing system or product (e.g. cows mixed) | Data represents majority of foods within a system
3 | Need to gapfill countries with considerable production (e.g. Philippines) | National | Data range doesn't include 2017 and is likely to represent 2017. | Generalized to animal type, but not rearing system or product (e.g. cows) | Data represents half of foods within a system
2 | Need to gapfill whole continents/georegions | Georegions | Data range includes 2017 and is unlikely to represent 2017 | Generalized to multiple animal types (e.g. cattle/ungulates) | Data represents minority of foods within a system
1 | Generalized globally using site-level data. | Global/generalized | Data is older than 2017 and is unlikely to represent 2017 | Data is generalized across major food system (e.g. all reared animals) | Data represents minimum of foods within a system
NA | Spatial extent is not relevant | Spatial resolution of the data is not relevant | Age of data isn't relevant | Data is not specific to a food system | Data is not specific to a food system

**Salmonids farms map quality**
 - Spatial extent: 4
 - Spatial resolution: 5
 - Temporal specificity: 4
 - System specificity: 2
 - System comprehensiveness: 4

**Tuna farms map quality**
 - Spatial extent: 5
 - Spatial resolution: 5
 - Temporal specificity: 4
 - System specificity: 3
 - System comprehensiveness: 5

**Shrimp farms map quality**
 - Spatial extent: 2
 - Spatial resolution: 5
 - Temporal specificity: 4
 - System specificity: 2
 - System comprehensiveness: 4

**General marine fish farms map quality**
 - Spatial extent: 2
 - Spatial resolution: 5
 - Temporal specificity: 4
 - System specificity: 1
 - System comprehensiveness: 4

**Unfed or algae fed shellfish farms map quality**
 - Spatial extent: 2
 - Spatial resolution: 5
 - Temporal specificity: 4
 - System specificity: 2
 - System comprehensiveness: 4

**Other crustaceans farms map quality**
 - Spatial extent: 1
 - Spatial resolution: 1
 - Temporal specificity: NA
 - System specificity: 1
 - System comprehensiveness: 4

Some random exploring
```{r}
by_type <- all_farms %>%
  group_by(species) %>%
  summarise(n_obs = n())
